---
output: html_document
---
<style>
html {
  font-family: Georgia;
  font-size: inherit;
}
body {
  font-family: inherit;
  font-size: inherit;
  font-variant-numeric: lining-nums;
}
h1 {
  font-size: 2rem;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<h1>Bagging for Regression Trees</h1>

Slide 11 - Apply bagging to the Boston data using the randomForest package in R
```{r echo=TRUE, collapse=TRUE, prompt=TRUE, comment=NA}
library(randomForest)
library(MASS)
set.seed(1)
train = sample(1:nrow(Boston), nrow(Boston)/2)
set.seed(6)
bag.boston = randomForest(medv~., data=Boston, subset=train, mtry=13, importance=TRUE)
bag.boston
```

Slide 12 - How well does this bagged model perform on the test set?
```{r echo=TRUE, collapse=TRUE, prompt=TRUE, comment=NA, fig.align="center"}
yhat.bag = predict(bag.boston, newdata=Boston[-train, ])
boston.test = Boston[-train, "medv"]
plot(yhat.bag, boston.test)
abline(0, 1)
mean((yhat.bag - boston.test)^2)
set.seed(6)
bag.boston = randomForest(medv ~ ., data=Boston, subset=train, mtry=13, ntree=25, importance=TRUE)
yhat.bag = predict(bag.boston, newdata=Boston[-train, ])
mean((yhat.bag-boston.test)^2)
```

Slide 14 - A comparison of Error Rates for Boston Housing Data
```{r echo=TRUE, collapse=TRUE, prompt=TRUE, comment=NA, fig.align="center"}
set.seed(1)
train = sample(1:nrow(Boston), nrow(Boston)*0.5)
train.df = Boston[train,]
test.df = Boston[-train,]
x_train = train.df[-14]
x_test = test.df[-14]
y_train = train.df$medv
y_test = test.df$medv
myForest1 = randomForest(x=x_train, y=y_train, xtest=x_test, ytest=y_test, ntree=100, mtry=13)
plot(1:100, myForest1$test$mse, main="Test Error from Random Forests on the Boston dataset", xlab="Number of Trees", ylab="Test MSE", type= "l", ylim=c(12,22), lwd=1)
abline(h=myForest1$test$mse[1], lty=2, col="red")
```

<h1>Bagging for Classification Trees</h1>

Slide 17 - Apply bagging to the Carseats data using the randomForest package in R
```{r echo=TRUE, collapse=TRUE, prompt=TRUE, comment=NA}
library(ISLR)
library(randomForest)
High = ifelse(Carseats$Sales <= 8, "No", "Yes")
Carseats = data.frame(Carseats, High)
set.seed(2)
train = sample(1:nrow(Carseats), nrow(Carseats)/2)
Carseats.test = Carseats[-train,]
High.test = Carseats[-train, "High"]
bag.carseats = randomForest(High ~ .-Sales, data=Carseats, subset=train, mtry=10)
yhat.carseats= predict(bag.carseats, newdata=Carseats.test)
table(yhat.carseats, High.test)
```

Slide 21 - Code to obtain the plot on [Slide 20]
```{r echo=TRUE, collapse=TRUE, prompt=TRUE, comment=NA, fig.align="center"}
library(ISLR)
library(randomForest)
data(Carseats)
High = ifelse(Carseats$Sales <= 8, "No", "Yes")
Carseats.mv = data.frame(Carseats, High)
Carseats.mv = Carseats.mv[, -1]
set.seed(2)
train = sample(400, 200)
Carseats.mv.test = Carseats.mv[-train]
High.mv.test = Carseats.mv[-train, 11]
```