---
output:
  html_document: default
---
<link href="https://fonts.googleapis.com/css?family=Cinzel" rel="stylesheet">
<style>
html {
  font-family: Georgia;
  font-size: inherit;
}
body {
  font-family: inherit;
  font-size: inherit;
  font-variant-numeric: lining-nums;
}
h1 {
  font-family: 'Cinzel';
  font-size: 2rem;
  font-weight: bold;
  color: #562435;
  text-align: center;
  margin: 1rem 0;
}
h2 {
  font-size: 1.2rem;
  font-weight: bold;
  color: #562435;
  margin-top: 2rem;
  border-top: 2px solid #562435;
}
h2 span {
  background-color: #562435;
  color: white;
  padding: 0 0.5rem 0.2rem;
}
div.question {
  font-weight: bold;
  color: #562435;
  margin-bottom: 1rem;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, collapse=TRUE, prompt=FALSE, comment=NA, fig.align="center")
```

<h1>Lab 7 - Bagging and Random Forests</h1>

<h2><span>Question 1</span></h2>

<div class="question">`survived` is a numeric value.&nbsp;&nbsp;We need to first transform it to a categorical value and save it as a new variable `survived01`.&nbsp;&nbsp;Use `titanic3$survived01 = as.factor(titanic3$survived)` to do so and check that this variable has been included in the dataset.</div>

```{r}
library(dplyr)
t3 = read.csv("titanic3")
t3 = select(t3, -name, -ticket, -boat, -body, -home.dest, -cabin)
t3$survived01 = as.factor(t3$survived)
head(t3)
```

<h2><span>Question 2</span></h2>

<div class="question">Install the package `randomForest` and include this in your code.&nbsp;&nbsp;In order to call the `randomForest()` function, all the missing value rows need to be dealt with.&nbsp;&nbsp;The simplest way is to remove those rows.&nbsp;&nbsp;Use `titanic3 = na.omit(titanic3)` to do that.</div>

```{r}
library(randomForest)
t3 = na.omit(t3)
```

<h2><span>Question 3</span></h2>

<div class="question">Use a seed to set half of the dataset to be the training dataset and the other half to be the test dataset.</div>

```{r}
set.seed(8)
train = sample(nrow(t3), nrow(t3)/2)
test = t3[-train, ]
x_test = test[, -c(2, 9)]
survived.test = t3$survived[-train]
survived01.test = t3$survived01[-train]
```

<h2><span>Question 4</span></h2>

<div class="question">Use the training dataset to build a bagged model for:</br>
&nbsp;&nbsp;&nbsp;&nbsp;y: `survived`</br>
&nbsp;&nbsp;&nbsp;&nbsp;x: all the features other than `survived` and `survived01`</br>
Compare the mean error rate on the test dataset.</div>

```{r}
set.seed(6)
t3.bag.reg = randomForest(survived ~ .-survived01, data=t3, subset=train, mtry=7, importance=TRUE)
t3.bag.reg
t3.pred.reg = predict(t3.bag.reg, newdata=test)
t3.pred.cls = ifelse(t3.pred.reg <= 0.5, "0", "1")
mean(t3.pred.cls != survived.test)
```

<h2><span>Question 5</span></h2>

<div class="question">Using the same training and test datasets, build a bagged model for:<br/>
&nbsp;&nbsp;&nbsp;&nbsp;y: `survived01`<br/>
&nbsp;&nbsp;&nbsp;&nbsp;x: all the features other than `survived` and `survived01`<br/>
a\. Find out how many trees your model has built and the OOB error.<br/>
b\. Compute the mean error rate on the training dataset.<br/></div>

```{r}
t3.bag.cls = randomForest(survived01 ~ .-survived, data=t3, subset=train, mtry=7, importance=TRUE)
t3.bag.cls
(t3.bag.cls$confusion[2,1] + t3.bag.cls$confusion[1,2]) / t3.bag.cls$ntree

```

<h2><span>Question 6</span></h2>

<div class="question">Plot the variable importance plot for the two bagged models you built in 4) and 5) and comment whether the importance coincides.</div>

```{r}
importance(t3.bag.reg)
varImpPlot(t3.bag.reg)
importance(t3.bag.cls)
varImpPlot(t3.bag.cls)
```

<h2><span>Question 7</span></h2>

<div class="question">Plot a graph that shows the test error rate of a single tree (red dashed line), the mean test error rate for majority vote (black curve) and the test error rates for averaging the probabilities (blue curve), both in relation to the number of trees.&nbsp;&nbsp;Add a legend if you can.</div>

```{r warning=FALSE}
set.seed(1)
index.train = sample(nrow(t3), nrow(t3)/2)
dataset.training = t3[train, ]
dataset.test = t3[-train, ]
y.actual = dataset.test[, 9]
number.of.iterations = 100
test.error.rate.mv = rep(0, number.of.iterations)
test.error.rate.av = rep(0, number.of.iterations)
for (i in 1:number.of.iterations) {
  # Majority vote:
  set.seed(6)
  rf.fit.mv = randomForest(survived01 ~ .-survived, data=dataset.training, mtry=7, ntree=i, importance=TRUE)
  y.predict.mv = predict(rf.fit.mv, newdata=dataset.test)
  confuse.mv = table(y.predict.mv, y.actual)
  test.error.rate.mv[i] = (confuse.mv[1,2] + confuse.mv[2,1]) / nrow(dataset.test)
  # Averaging Probabilities:
  set.seed(6)
  rf.fit.av = randomForest(survived ~ .-survived01, data=dataset.training, mtry=7, ntree=i, imporance=TRUE)
  y.predict.av = ifelse(predict(rf.fit.av, newdata=dataset.test) <= 0.5, "0", "1")
  confuse.av = table(y.predict.av, y.actual)
  test.error.rate.av[i] = (confuse.av[1,2] + confuse.av[2,1]) / nrow(dataset.test)
}
plot(test.error.rate.av,
  col="blue",
  type="l",
  xlab="Number of Trees in Forest",
  ylab="Error Rate (Test Dataset)",
  main="Error Rates for Different Sized Forests",
  ylim=c(0.19,0.30)
)
lines(test.error.rate.mv,
  col="black",
  type="l"
)
abline(h=test.error.rate.mv[1], col="red", lty=2)
legend("topright",
  c("single tree", "majority vote", "averaging probabilities"),
  col=c("red", "black", "blue"),
  lty=c(2, 1, 1)
)
```

<h2><span>Question 8</span></h2>

<div class="question">Plot a graph tht shows the best value of mtry for the random forest model
&nbsp;&nbsp;&nbsp;&nbsp;y: `survived01`<br/>
&nbsp;&nbsp;&nbsp;&nbsp;x: all the features other than `survived` and `survived01`<br/>
&nbsp;&nbsp;&nbsp;&nbsp;mtry: range from 1 to 7<br/>
</div>

```{r warning=FALSE}
m = 7
test.error.rate = rep(0, m)
for (i in 1:m) {
  set.seed(6)
  rf.fit = randomForest(survived ~ .-survived01, data=dataset.training, mtry=i, imporance=TRUE)
  y.predict = ifelse(predict(rf.fit, newdata=dataset.test) <= 0.5, "0", "1")
  confuse = table(y.predict, y.actual)
  test.error.rate[i] = (confuse[1,2] + confuse[2,1]) / nrow(dataset.test)
}
plot(test.error.rate,
  col="blue",
  type="b",
  xlab="mtry",
  ylab="Error Rate (Test Dataset)",
  main="Averaging Probability Error Rates for Different 'mtry' values"
)
```

<h2><span>Question 9</span></h2>

<div class="question">Play with mtry and ntree, plot a graph that shows test error rate vs. ntree for different mtry and find the best/reasonably good combination of mtry and ntree from the plot.&nbsp;&nbsp;Add a legend if you can.</div>

```{r}

```